OneD: increasing reproducibility of Hi-C Samples with abnormal karyotypes

### 2017-06-06 main.tex
Wrote the section about copy number variation.

\begin{equation*}
\hat{x}_{ij}^* = \frac{\hat{x}_{ij}}{\sqrt{c_ic_j}}
\end{equation*}

@@

@@

\begin{figure}
\centerline{\includegraphics[width=.5\textwidth]{img/times_figure5.pdf}}
\caption{
Computing time of the bias correction methods. A. Total time for the entire genome at 100 kbp resolution. Each dot corresponds to
one sample. The only method faster than ours under performs in all sample
comparisons. B. Time to correct a reduced genome (chr19-22) of one sample at different resolutions. Note the logarithmic scale on the y-axis on both panels.}
\label{fig:times}
\end{figure}

>>>>

\begin{equation}
\label{eq:cnvnorm}
\hat{x}_{ij}^* = \frac{\hat{x}_{ij}}{\sqrt{c_ic_j}}
\end{equation}

@@

Removing the explicit experimental biases enhancers the importance of the
copy number in the signal (Figure~\ref{fig:copy_number}). The copy number
is not an experimental bias, but it may be considered an additional source
of bias to be remove. For this reason, OneD also allows the user to
correct the copy number and output a matrix equivalent to that of a normal
karyotype.

For this application, OneD segments the linear signal of the total amount
of contacts into piecewise homoegenous regions. This is carried out by a
hidden Markov model whereby the averages of the states are constrained to
be within a discrete factor of each other. This allows the model to detect
regions with a number of copies equal to 1, 2, 3 \textit{etc}. With the
inferred number of copies at hand, OneD then normalizes each cell of the
matrix with equation (\ref{eq:cnvnorm}).

Figure~\ref{fig:cnv_correction} shows the effect of this process on the
final Hi-C matrix. Panel A indicates that different normalization methods
will either enhancer or diminish the signal in regions with higher copy
number. In this example, ICE overcompensates the original bias at the
center of the picture and fades the signal alost entirely. Concomitantly,
the signal at the bottom left of the matrix is enhanced to make appear a
structure not visible in the raw data. On the contrary, \textit{lgf}
strengthen the central region and the diagonal. OneD reduces the level of
the central portion by a factor 2 approximately, but does otherwise not
distort the main features of the region. This example shows that the copy
number does not have a simple and predictable effect on the final matrix.
Not taking it into account may open the door to some normalization
artifacts. The result provided by OneD is not necessarily the right one
(see Discussion), but at least it does not compensate copy number
variations as a side effect of some other criteria.

\begin{figure}
\centerline{\includegraphics[width=.5\textwidth]
  {img/figure_cnv_correction.pdf}}
\caption{
Copy number correction A. Detail of a Hi-C matrix normalized with
different methods. The central portion has an increased copy number, which
affects normalization. ICE fades it away, lgf enhances it and OneD reduces
the signal by about half. B. Profile of the total amount of contacts after
copy numer correction. The plot shows the same region as
Figure\ref{fig:totals}B. The brown line represents the raw signal, the
blue line represents the signal after bias correction, the black line
represents the signal after bias and copy number correction.}
\label{fig:cnv_correction}
\end{figure}

Figure~\ref{fig:cnv_correction}B shows the total sum of contacts after
correction for the copy number with OneD. The signal that originally
showed a sharp transition corresponding to the gain or loss of a large
region, now shows a flat trend throughout the chromosome.

In summary, OneD can be used to obtain a normalized matrix equivalent to
that of a normal karyotype in cases that the effect of the copy number
must be removed from the signal.

@@

\begin{figure}
\centerline{\includegraphics[width=.5\textwidth]
  {img/figure_benchmark_time.pdf}}
\caption{
Computing time of the bias correction methods. A. Total time for the entire genome at 100 kbp resolution. Each dot corresponds to
one sample. The only method faster than ours under performs in all sample
comparisons. B. Time to correct a reduced genome (chr19-22) of one sample at different resolutions. Note the logarithmic scale on the y-axis on both panels.}
\label{fig:times}
\end{figure}
